Weak interspecific interactions in a sagebrush steppe: Evidence from observational data, population models, and a removal experiment

### 2016-05-23 removals/manuscript/removal_ms.tex
tiny changes

\author[a]{Peter B. Adler\thanks{Corresponding author. Department of Wildland Resources and the Ecology Center, Utah State University, Logan, Utah Email: peter.adler@usu.edu}}
\author[a]{Andrew Kleinhesselink}
\author[b]{Stephen P. Ellner}
\author[b]{Giles Hooker}
\author[a]{Brittany Teller}
\affil[a]{Department of Wildland Resources and the Ecology Center, Utah State University, Logan, Utah 84322}
\affil[b]{Department of Ecology and Evolutionary Biology, Cornell University, Ithaca, New York}

@@

\begin{enumerate}
\item. A previous study of four dominant species from a sagebrush steppe community used historical data sets and multispecies population models to demonstrate weak interspecific interactions, implying very little competitive release. 
\item. We conducted a removal experiment to test the model's ability to predict competitive release. We established new quadrats in the same location where the historical data were collected. We assigned half the new quadrats to a  perennial grass removal treatment and half to a treatment in which we removed the dominant shrub, \textit{Artemisia tripartita}.
\item. We modeled survival, growth and recruitment as functions of local neighborhood species composition along with an indicator variable for removal treatment. For survival and recruitment, the removal treatment effects were never significantly positive, indicating that competitive release was not underestimated by the interspecific effects included in the ``baseline'' model. 
\item. For growth, the removal treatment effect was significant and positive for two species, \textit{Poa secunda} and \textit{Pseudoroegneria spicata}, indicating that the baseline model underestimated competitive release. However, including information about the location of these grass individuals with respect to removed \textit{Artemisia} failed to improve the growth regressions, raising questions about the mechanisms driving the positive response to removal.
\item. For three of the four species, individual based models and integral projections models  based on the vital rate regressions showed that removal treatment effects had little impact on population growth rates or simulated equilibrium cover following a species removal. For \textit{P. spicata}, the population models that included effects of \textit{Artemisia} removal projected much stronger competitive release than the baseline models. 
\item. \emph{Synthesis}: Although models based on observational data did not perfectly predict species responses to removal, overall the results increase our confidence that competitive release is in fact small for at least three of our four study species. The implication is that resource competition is not the primary factor determining the dynamics of this community.

>>>>

\author[a]{Peter B. Adler\thanks{Corresponding author. Department of Wildland Resources and the Ecology Center, Utah State University, Logan, Utah Email: peter.adler@usu.edu}}
\author[a]{Andrew Kleinhesselink}
\author[b]{Stephen P. Ellner}
\author[c]{Giles Hooker}
\author[a]{Brittany Teller, and others? Robin? Bret?}
\affil[a]{Department of Wildland Resources and the Ecology Center, Utah State University, Logan, Utah 84322}
\affil[b]{Department of Ecology and Evolutionary Biology, Cornell University, Ithaca, New York}
\affil[c]{Department of Biological Statistics and Computational Biology, Cornell University, Ithaca, New York}

@@

\begin{enumerate}
\item A previous study of four dominant species from a sagebrush steppe community used historical data sets and multispecies population models to demonstrate weak interspecific interactions, implying very little competitive release. 
\item We conducted a removal experiment to test the model's ability to predict competitive release. We established new quadrats in the same location where the historical data were collected. We assigned half the new quadrats to a  perennial grass removal treatment and half to a treatment in which we removed the dominant shrub, \textit{Artemisia tripartita}.
\item We modeled survival, growth and recruitment as functions of local neighborhood species composition along with an indicator variable for removal treatment. For survival and recruitment, the removal treatment effects were never significantly positive, indicating that competitive release was not underestimated by the interspecific effects included in the ``baseline'' model. 
\item For growth, the removal treatment effect was significant and positive for two species, \textit{Poa secunda} and \textit{Pseudoroegneria spicata}, indicating that the baseline model underestimated competitive release. However, including information about the location of these grass individuals with respect to removed \textit{Artemisia} failed to improve the growth regressions, raising questions about the mechanisms driving the positive response to removal.
\item For three of the four species, individual based models and integral projections models  based on the vital rate regressions showed that removal treatment effects had little impact on population growth rates or simulated equilibrium cover following a species removal. For \textit{P. spicata}, the population models that included effects of \textit{Artemisia} removal projected much stronger competitive release than the baseline models. 
\item \emph{Synthesis}: Although models based on observational data did not perfectly predict species responses to removal, overall the results increase our confidence that competitive release is in fact small for at least three of our four study species. The implication is that resource competition is not the primary factor determining the dynamics of this community.
